---
title: "Fishery Revenue Model"
author: "Maegan Blansett and Jeff Cedarbaum"
date: "May 24, 2017"
output: html_document
---


## Fishery Revenue Model 

The following model has been created to return information regarding total catch and revenue amounts for salmon fisheries in different locations along the West Coast. 

It requires the use of the following packages: 

* roxygen2

* ggplot2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```


The model, with associated **documentation**, is provided below. 


```{r fish model}
#' Fishery Revenue 
#' 
#' This function calculates the total revenue for different catches and prices of fish at different fishery locations. 
#' 
#' @param fish_price a table that gives the prices for different types of fish 
#' @param fish_catch a table that gives the number of fish caught at each location, where each location is a different column and each type of fish is a different row 
#' @return most frequently caught fish at each location 
#' @return total revenue at each location 
#' @return total revenue at all locations 
#' @return graph of revenue by location 


fishery_revenue = function(fish_price, fish_catch) {
  
  # return most frequently caught fish at each location 
  freq <- list(rownames(fish_catch[apply(fish_catch, 2, which.max), ]), colnames(fish_catch[apply(fish_catch, 2, which.max), ]))
  
  # return total revenue at each location 
  new_fish <- (fish_catch) * (fish_price$price)
  loc_rev <- apply(new_fish, 2, sum)
  
  # return total revenue at all locations 
  all_rev <- all_rev <- sum(loc_rev)
  
  # if user desired, graph revenue by location 
  plot_data <- as.data.frame(cbind(location, loc_rev)) #Creating dataframe to plot
  
  scaleFUN <- function(x) sprintf("%.2f", x) #limiting decimal places on y-axis
  
  rev_plot <- ggplot(plot_data)+
    geom_bar(aes(x=location, y=as.numeric(as.character(loc_rev))), stat="summary", fun.y="sum", col="Blue", fill="darkgreen")+
    labs(x="Location", y="Revenue (million $)")+
    ggtitle("Salmon Fishing Revenue by Location")+
    scale_y_continuous(labels=scaleFUN)
    # geom_text(aes(label=as.numeric(as.character(loc_rev), digits = 2), x=location, y=as.numeric(as.character(loc_rev)), vjust=1.5)) ##Adding this line of code will insert text onto the plot showing exact revenue for each location, if the user desires.

  
  return(list(freq, loc_rev, all_rev, rev_plot))
  
}
```

## Generating Sample Data  

In order to successfully utilize this model, the user must ensure that their data are provided in the correct format. 

Required inputs are: 

* A data frame called **fish_price**, which contains a column of fish types as a *factor* and a column of their respective current market prices ($USD per fish) as a *numeric*. 

* A data frame called **fish_catch**, which contains a column for each fishery location, a row for each type of fish, and counts of fish caught at each location. 

The following code can be used to generate a **fish_price** input table. 

```{r fish_price}
# Create a table of fish prices for different types of salmon:

# First, create vector of types of salmon
fish <- c("chinook", "coho", "pink", "sockeye", "chum", "steelhead", "cutthroat" )

# Next, create a vector of prices
price <- c(375, 300, 250, 350, 260, 340, 355)
price
class(price) # numeric 

# Finally, bind the fish names and prices into one table 
fish_price <- data.frame(cbind(fish, price))
class(fish_price$price) # now changed to factor
fish_price$price <- as.numeric(paste(fish_price$price)) # paste makes the values stay what they are
View(fish_price)
```

The following code can be used to generate a **fish_catch** input table. 

```{r fish_catch}
# Create a table of fish catches at different locations:

# First, create a vector of possible locations 
location <- c("Anchorage", "Juneau", "Prince Rupert", "Vancouver", "Seattle", "Portland", "Klamath", "Eureka")

# Next, create an empty data frame of fish types and locations
fish_catch <- data.frame(matrix(ncol = length(location), nrow = length(fish)))
rownames(fish_catch) = fish
colnames(fish_catch) = location
View(fish_catch)

# Finally, fill in the values of the table with random numbers of fish caught 
fish_catch$Vancouver <- rnorm(6, sd = 0.5, n = length(fish))
fish_catch$Seattle <- rnorm(6, sd = 0.5, n = length(fish))
fish_catch$Portland <- rnorm(6, sd = 0.5, n = length(fish))
fish_catch$Klamath <- rnorm(6, sd = 0.5, n = length(fish))
fish_catch$Juneau <- rnorm(6, sd = 0.5, n = length(fish))
fish_catch$Anchorage <- rnorm(6, sd = 0.5, n = length(fish))
fish_catch$`Prince Rupert` <- rnorm(6, sd = 0.5, n = length(fish))
fish_catch$Eureka <- rnorm(6, sd = 0.5, n = length(fish))
View(fish_catch)
```

## Running the Model

To run the model, simply use the following code. See results below.

```{r model}
fishery_revenue(fish_price, fish_catch)
```

