---
title: "Informatics HW3"
author: "Jeff Cedarbaum"
date: "May 18, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Before beginning I need to import the data that I'll be visualizing, and make sure the proper R packages are loaded.

```{r data import}

setwd("c:/Users/j/Desktop/planets/ggplot2")

thinshallow <- read.table("data/resthin.shallow.txt", header=T)
thindeep <- read.table("data/resthin.deep.txt", header=T)
clim <- read.table("data/sierraclim.txt",header=T)

library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
library(lubridate)
library(tidyverse)
```

Now that data and packages are loaded, first I'll take a look at differences in the rate of evapo-transpiration in "shared" vs "not shared" thinning practices in the thindeep dataset.

```{r thindeep, echo = TRUE}

thindeep$ET <- thindeep$trans+thindeep$evap

shared <- subset(thindeep, thindeep$shared=="TRUE")
unshared <- subset(thindeep, thindeep$shared=="FALSE")

Plot_shared <- ggplot(shared)+
  geom_line(aes(x = wy, y = ET), stat="summary", fun.y="mean", col = "Green", size = 1)+
  ggtitle("Shared Thinning")+
  labs(x = "Years Since Thinning", y = "Evapo-Transpiration", size=8)

Plot_unshared <- ggplot(unshared)+
  geom_line(aes(x = wy, y = ET), stat="summary", fun.y="mean", col = "Blue", size = 1)+
  ggtitle("Unshared Thinning")+
  labs(x = "Years Since Thinning", y = "Evapo-Transpiration", size=8)

grid.arrange(Plot_shared,Plot_unshared)

```


Next, just using the 'shared' data points from the thindeep dataset, I'll explore the relationship between evaporation and transpiration.

```{r E&T, echo = TRUE}

Plot_ET <- ggplot(shared)+
  geom_line(aes(x=month, y=evap), stat="summary", fun.y = "mean", col = "Blue")+
  geom_line(aes(x=month, y=trans), stat="summary", fun.y = "mean", col = "Green")+
  labs(x="Month", y="Water Evaporated or Transpired")+
  annotate("text", y=0.25, x=6, label = "Evaporation")+
  annotate("text", y=3, x=6, label = "Transpiration")+
  ggtitle("Transpiration and Evaporation under Shared Thinning")+
  annotate("text", y=-0.3, x=1, label = "Jan")+
  annotate("text", y=-0.3, x=2, label = "Feb")+
  annotate("text", y=-0.3, x=3, label = "Mar")+
  annotate("text", y=-0.3, x=4, label = "Apr")+
  annotate("text", y=-0.3, x=5, label = "May")+
  annotate("text", y=-0.3, x=6, label = "Jun")+
  annotate("text", y=-0.3, x=7, label = "Jul")+
  annotate("text", y=-0.3, x=8, label = "Aug")+
  annotate("text", y=-0.3, x=9, label = "Sep")+
  annotate("text", y=-0.3, x=10, label = "Oct")+
  annotate("text", y=-0.3, x=11, label = "Nov")+
  annotate("text", y=-0.3, x=12, label = "Dec")

Plot_ET


  
  #Plot_ET <- ggplot(shared)+
#  geom_line(aes(x=month, y=as.factor(evap)), stat="summary", fun.y = "mean", col = "Blue")+
#  geom_line(aes(x=month, y=as.factor(trans)), stat="summary", fun.y = "mean", col = "Green")+
#  labs(x="Month", y="Water Evaporated or Transpired")+
#  annotate("text", y=0.25, x=6, label = "Evaporation")+
#  annotate("text", y=3, x=6, label = "Transpiration")+
#  ggtitle("Transpiration and Evaporation under Shared Thinning")
#  scale_x_discrete(labels=month.abb)

  #I tried this ^ code to get the x-axis labels to be month names instead of numbers, I'm not sure why it didn't work.

```


For the next one, I'll split average temperature by month using the clim dataset.

```{r temp, echo = TRUE}

clim$ave_temp <- (clim$tmin+clim$tmax)/2

plot_monthly_temp <- ggplot(clim)+
  geom_bar(aes(x=month, y = ave_temp), stat="summary", fun.y="mean", col="red", fill="red")+
  labs(x="Month", y="Average Temperature (degrees C)")+
  ggtitle("Average Monthly Temperature")+
  annotate("text", y=-0.3, x=1, label = "Jan")+
  annotate("text", y=-0.3, x=2, label = "Feb")+
  annotate("text", y=-0.3, x=3, label = "Mar")+
  annotate("text", y=-0.3, x=4, label = "Apr")+
  annotate("text", y=-0.3, x=5, label = "May")+
  annotate("text", y=-0.3, x=6, label = "Jun")+
  annotate("text", y=-0.3, x=7, label = "Jul")+
  annotate("text", y=-0.3, x=8, label = "Aug")+
  annotate("text", y=-0.3, x=9, label = "Sep")+
  annotate("text", y=-0.3, x=10, label = "Oct")+
  annotate("text", y=-0.3, x=11, label = "Nov")+
  annotate("text", y=-0.3, x=12, label = "Dec")


plot_monthly_temp

```



Finally, I'll look at average temperatures during summer (June through September) vs the rest of the year, side by side.


```{r last plot, echo = TRUE}


clim$summer <- ifelse(clim$month>5 & clim$month<10, "Summer", "Not Summer")


plot_temp_facet <- ggplot(clim)+
  geom_point(aes(x=year, y = ave_temp), stat="summary", fun.y="mean", col = "red")+
  labs(x="Year", y="Average Temperature (degrees C)")+
  ggtitle("Average Annual Temperature")+
  facet_wrap(~as.factor(clim$summer))

#Tried to use colorbrewer to define color by average temperature and get a diverging palette. It didn't work and I'm not sure why. 

plot_temp_facet

```




